\documentclass[runningheads]{llncs}

\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{textcomp}
\usepackage{float}
\usepackage{tabularx}
\usepackage{bm}
\usepackage{enumitem}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{listings}
\usepackage{minted}
\usepackage{caption}
\usepackage[hidelinks]{hyperref}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{seqsplit}
\usepackage[breaklinks]{hyperref}
\usepackage[hyphens]{url} 
\usepackage{xurl}

\lstset{
  breaklines=true,
  breakatwhitespace=false,
  columns=fullflexible,
  keepspaces=true,
  showstringspaces=false,
  basicstyle=\ttfamily\footnotesize
}

\captionsetup{compatibility=false}
\urlstyle{tt}
\floatname{listing}{Code}
\usemintedstyle{friendly}
\newcommand{\ttbreak}[1]{\texttt{\seqsplit{#1}}}
\setlength{\parskip}{0pt}
\setlength{\parindent}{1em}
\setlength{\textfloatsep}{8pt}
\setlength{\intextsep}{8pt}
\setlength{\floatsep}{8pt}
\raggedbottom

\begin{document}

\title{MeReader: An Offline, Privacy-Preserving, Progress-Aware AI Assistant for Narrative eBook Reading}
\author{Mohammed Efaz\inst{1}\orcidID{0009-0002-3562-7607} \and
Itilekha Podder\inst{1}\orcidID{0000-0001-7999-5597} \and
Udo Bub\inst{1}\orcidID{0000-0002-6018-2411}}
\authorrunning{M. Efaz et al.}
\institute{Faculty of Informatics, Eötvös Loránd University (ELTE), Budapest, Hungary\\
\email{efaz@mohammedefaz.com, itilekha19@inf.elte.hu, udobub@inf.elte.hu}}
\maketitle

\begin{abstract}
This paper introduces MeReader, a privacy-perserving, offline and progress-aware AI assistant designed to function as a contextual companion for narrative reading on consumer hardware.
Unlike conventional Large Language Model (LLM)-based systems that process documents/books online in an all-texts-at-once type of way, thereby risking narrative spoilers, MeReader confines both retrieval and generation to the reader's distinct position in the text.
The system integrates a hybrid retrieval architecture, combining vector similarity, BM25, and summary embeddings, to ensure high fidelity to the reader's current context.
We evaluate the system across classical English books using both quantitative comprehension metrics and qualitative LLM-based independent judging.
The results identify a quality-latency frontier, showing that mid-sized local models (2b-4b) can achieve comprehension parity with larger counterparts (5b+) while maintaining acceptable response quality.
Importantly, the proposed progress-aware boundary mechanism is shown to prevent future content leakage without compromising retrieval quality and correctness.
These findings suggest that a localized, privacy-centric approach can augment, rather than replace, the digital reading experience, offering a good alternative to cloud-dependent services. 
Therein, there remains a lack of amalgamate systems which combine progress-aware, privacy-preserving, and narrative eBook reading with AI support.
It is in this very gap that the development of MeReader came to be; synthesizing said strands into an offline, book-constrained assistant focused on enhancing comprehension
without compromising narrative continuity or user privacy.
\keywords{Progress-aware retrieval \and RAG \and On-device LLM \and eBook reader \and Privacy-preserving.}
\end{abstract}

\section{Introduction}
\label{sec:introduction}
\input{chapters/01_intro}
\section{Related Work}
\label{sec:related_work}
\input{chapters/02_Related_Work}
\section{Proposed System Design}
\label{sec:system_design}
\input{chapters/03_System_Design}
\section{Experimental Setup}
\label{sec:implement}
\input{chapters/04_Implementation}
\section{Results and Discussion}
\label{sec:evaluation}
\input{chapters/05_Evaluation}
\section{Conclusion and Future Work}
\label{sec:conclusion}
\input{chapters/06_Conclusion}

\bibliographystyle{splncs04}
\bibliography{references}

\end{document}